Bayes Optimal Informer Sets for Drug Discovery\thanks{abc}

### 2020-04-15 BOISE/Paper_draft/draft.tex
draft update

\usepackage{amssymb,amsmath,amsfonts,bbm,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,caption,natbib,pdflscape,subfigure,array,hyperref}

\usepackage{algorithm, algorithmicx}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{appendix}

@@

\begin{titlepage}
\title{Bayes Optimal Informer Sets for Drug Discovery\thanks{abc}}
\author{Peng Yu, {\sc contributing members of UW drug discovery group}, \\  \& Michael A. Newton\thanks{abc}}
\date{\today}
\maketitle
\begin{abstract}
\noindent Placeholder\\
\vspace{0in}\\
\noindent\textbf{Keywords:} key1, key2, key3\\
\vspace{0in}\\
\noindent\textbf{JEL Codes:} key1, key2, key3\\

@@

\section{Background} \label{sec:background}

Zhang, Ericksen, and Lee {\em et al.} (2019) (hereafter, ZEL) introduced
informer-based-ranking (IBR) strategies for computational
guidance of drug-discovery experiments.  The context is 
virtual screening (Sliwoski {\em et al.} 2014).   Briefly, a chemical screening facility receives 
material representing a purified protein target, and is tasked with identifying drugs
that may have some inhibitory effect on the bioactivity of the protein.  For example, ZEL report
on drugs that affect certain protein kinases.
The facility has specific biochemical assays that test for
inhibitory effects of a given drug, and it has access to a very large library of different drug compounds, some small fraction of which are expected to have bioactivity
against the new target.   
Costs limit the number of compounds that may be screened experimentally.   The virtual screening task is to prioritize the drugs 
in advance of experimental screening by some computations so as to
improve the hit rate of the experiments performed on the new target.

Virtual screening may operate on different kinds of target and compound data, including
possibly chemical structures or the results of experimental assays.   For example, when chemical structures are available
it is possible to use sophisticated molecular docking programs that aim to find best fits between target and compound
structures (e.g. cite papers, including our own consensus docking paper.)
We focus on IBR strategies that have access to data on the bioactivity of many compounds
against many targets, though possibly not any bioactivity data from 
the new target on test. **note...no compound structure data is needed**  With these initial bioactivity data, IBR  
identifies a subset of compounds -- the {\em informer set} -- which
represent a small fraction of the available compounds and which we
will test experimentally for bioactivity against the new target.  Subsequently, we use the full set of data (initial data on other targets
for many compounds and intermediate data on informer compounds on 
the new target) in order to prioritize the remaining, non-informer
compounds for their likely bioactivity against the new target.  If
we have done a good job in building the informer set, then the 
top of the prioritized list of non-informers will be enriched
for truly active compounds.  We will have spent computational time
building the informer set from initial data, and experimental resources
obtaining intermediate data; the payoff is that further experiments
will return truly active drug compounds at  a higher rate than
if we had selected compounds to test by some other scheme.

ZEL introduced several bioactivity-based IBR strategies and compared
them empirically in data on protein kinases. Though differing in detail,
the proposed strategies all partition the initial
bioactivity data, producing  clusters of relatively similar targets; 
then informer compounds are  those that are predictive of the cluster label of a target.  Among kinase targets, the proposed bioactivity-based strategies had better predictive
performance than commonly used strategies based upon chemical
structure.   Since effective IBR strategies leverage statistical
patterns in the initial bioactivity data, we reason that
statistical modeling may provide a useful approach to deriving
more effective IBR strategies than are currently available. 
In the present work, we investigate a decision-theoretic formalism for constructing informer sets ***

\section{Bayes Optimal Informer SEt (BOISE)}

@@

We are given an $m \times n$ binary matrix $x_0$ containing bioactivity data measured on $m$ targets with index set $I = \{1,2,\cdots,m\}$, and $n$ compounds with index set $J=\{1,2,\cdots,n\}$.  Each entry $x_{ij}$ of $x_0$ is a summary of experiments, such as inhibition assays,
of compound $j$ on target $i$. To be more specific, $x_{ij} = 1$ means compound $j$ is active on target $i$ while $x_{ij} = 0$ means inactive. Binary outcome is of more interest to us because our task is to determine and find certain compounds that are effective on the given target. It may worth to say that although our method is based on binary matrix, it can be applied to any continuous bioactivity matrices as we can always convert a continuous matrix into a binary matrix. For example, if $z$ is a continuous matrix, we can use a target-wise threshold criterion (Eq. \ref{eq:thresh}) to binarize $z$ based on sample mean and standard deviation of each row:

\begin{equation}\label{eq:thresh}
x_{ij} = \begin{cases}
		1, &\text{ if }
	      z_{ij} \geq \mbox{mean}[z_{i\cdot}] + 2 \times \mbox{s.d.}[z_{i\cdot}],\\
		0 & \text{ if }
	      z_{ij} < \mbox{mean}[z_{i\cdot}] + 2 \times \mbox{s.d.}[z_{i\cdot}].
	\end{cases}
\end{equation}
Throughout the paper we will use (Eq. \ref{eq:thresh}) to binarize target outcomes.

@@

In Zhang, Ericksen, and Lee {\em et al.} (2019), there are $3$ novel informer-based-ranking (IBR) methods: Regression Selection (RS), Coding Selection (CS), Adaptive Selection (AS), as well as $2$ baseline methods: Baseline Chemometric (BC) and Baseline Frequent-hitters (BF). In this paper, we will introduce a new IBR method named as Bayes Optimal Informer SEt (BOISE) that is uniformly better than all the methods mentioned above.

@@

\textbf{Note:} The reason we use notation $\mathcal{C}$ is that we consider clustering structure as a random process on all possible partitions of countable targets space. Dirichlet process mixture model (DPMM) is a classical method to handle this kind of clustering structures, and we are using a special case of DPMM called Chinese restaurant (CR) process based on some prior mass
$m_0$ to make the posterior inference. There are two main advantages of using DPMM instead of traditional clustering methods like spectral clustering, hierarchical clustering and k-means: On the one hand, we don't need to specify the number of clusters $K$ beforehand, or manually cut the tree to get the clustering structure. All we need to set is a prior mass $m_0$ that describes our belief of divergence and all the rest is data-driven. Thus, DPMM is less subjective than other methods. On the other hand, DPMM is a non-parametric Bayesian clustering method and hence is compatible with our Bayesian framework. Also, since the sample size is not very large($\approx 200$ targets), number of targets falling in each cluster can be unbalanced, with some clusters containing more than $50$ targets while other clusters containing less than 5 targets. DPMM is well designed to address this unbalanced impact under Bayesian framework. We will come to details of DPMM later in Section $3$ computation.

@@

We are faced with a two-stage decision problem: First, we need a rule to produce an informer set $A$ of size $n_A$ from $x_0$. Then we need to produce a top set $T$ of size $n_T$ based on initial data $x_0$, informer set $A$ and intermediate data $x_A$. Denote these decision rules as $A = A(x_0)$ and $T = T(A,x_A,x_0)$ to show their dependency.

@@

Consider when we have initial data $x_0$ as well as intermediate data $x_A$ on
some pre-specified informer set $A$, we can get the posterior distribution of $\theta$:
\begin{eqnarray*}
p(\theta|x_0,x_A) = \frac{ p(x_0|\theta) p(x_A|x_0, \theta) p(\theta) }{ p(x_0, x_A) }.
\end{eqnarray*}
For all compounds $j$, we then construct the posterior means under this distribution as:
\begin{eqnarray*}
\hat \theta_{i^*j} = \mathbb{E}\left( \theta_{i^*j} |x_A, x_0 \right)
\end{eqnarray*}
We define a particular top set rule $T^*$ to select the top $n_T$ compounds with largest posterior mean:
\begin{eqnarray}
\label{eq:topset}
T^*(A,x_A,x_0) = \left\{ j \in J: \; \sum_{k\in J} \mathbbm{1}(\hat \theta_{i^*k} > \hat \theta_{i^*j}) \leq n_T \right\}.
\end{eqnarray}
For the loss~(\ref{eq:loss}), it is known that $T^*$ minimizes the Bayes risk (Henderson, 2015, Proposition 2.3.1).
We utilize this finding to resolve one part of our compound-decision problem, obtaining:

@@

The first method is a straightforward sequential selection. At first we evaluate ${\rm PEL}_1(x_0,A)$ for all $A=1,\cdots,n$ to select the best $A$ with least ${\rm PEL}_1$ loss. Then at each iteration we memorize all previous decisions $A$ and select one best element $j\in J\setminus A$ that minimizes ${\rm PEL}_1(A\cup \{j\})$. The procedure can be summarized as following Algorithm \ref{Algo3}:
\begin{algorithm}
\caption{Exact Informer Selection}\label{Algo3}
\hspace*{\algorithmicindent} \textbf{Input:} Initial data $x_0$, size of informer set $n_A$, size of top set $n_T$.\\
\hspace*{\algorithmicindent} \textbf{Output:} Selected informer set of length $n_A$
\begin{algorithmic}[1]
\State \textbf{Initialization}: Evaluate ${\rm PEL}_1(x_0,A)$ for all $|A|=1$. Let $A^* = \argmin_{|A|=1}{\rm PEL}_1(x_0,A) $.
\While{$|A^*| < n_A$}
    \State Evaluate ${\rm PEL}_1(x_0,A^*\cup\{j\})$ for each $j\in J\setminus A^*$.
    \State Let $A^* \gets A^*\cup\{j^*\}$, where $j^* = \argmin_{j}{\rm PEL}_1(x_0,A^*\cup\{j\})$.
\EndWhile
\State Return $A^*$ as selected informer set.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Approximate Informer Selection}\label{Algo4}
\hspace*{\algorithmicindent} \textbf{Input:} Initial data $x_0$, size of informer set $n_A$, size of top set $n_T$.\\
\hspace*{\algorithmicindent} \textbf{Output:} Selected informer set of length $n_A$
\begin{algorithmic}[1]
\State \textbf{Initialization}: Let $A^* = \{\}$. Sample clustering assignments $\mathbb C$ from $p(\mathcal C\,|\, x_0)$ with Algorithm \ref{Algo1}.
\For{each $\mathcal C$ in $\mathbb C$}
    \State Sample $\mathbb{X}(\mathcal{C})$ from $p\left(x_{i^*}\,|\, \mathcal C,x_0\right)$ as in~(\ref{eq:Postphi}) and~(\ref{eq:Postc}).
    \Comment{\textbf{Auxiliary sample}}
\EndFor
\While{$|A^*| < n_A$}
    \For{each $j\in J\setminus A^*$}
        \For{each $\mathcal{C}$ in $\mathbb C$ and corresponding $\mathbb{X}(\mathcal{C})$}
            \State Evaluate $H(c(i^*)\,|\,\mathcal{C},x_0,x_{A^*\cup j})$ as in~(\ref{eq:Entropy});
        \EndFor
        \State Average over $x_{A^*\cup j}$ and $\mathcal{C}$ to get $\mathcal H(A^*\cup\{j\} )$ in~(\ref{eq:Score}). 
    \EndFor
    \State Let $A^*\gets A^*\cup \{j^*\}$, where $j^*$ is the minimizer of $\mathcal H(A^*\cup\{j\})$.
\EndWhile
\State Return $A^*$ as selected informer set.
\end{algorithmic}
\end{algorithm}
As mentioned in the beginning of this section, the major advantage of Algorithm \ref{Algo4} compared with Algorithm \ref{Algo3} is running time. In Algorithm \ref{Algo4} we bypass the computation of $\mathbb E\left(\theta_{ij}\,|\, x_0,x_A\right)$ and only use $p_k = P(i^*\in c_k\,|\, \mathcal C,x_0,x_A)$ to select informer set. This can improve our BOISE running time from days to hours while still maintaining similar performance. Therefore, we name it as "Accelerated-BOISE" in our empirical analysis. Also, although Accelerated-BOISE uses different informer selection method as BOISE does, they both share the same top set rule $T^*(A,x_A,x_0)$, which scores each $j\in J$ with their posterior expectation and choose the top $n_T$ as top set.

\subsection{Missing Data Imputation}
Previous informer based ranking methods in Zhang, Ericksen, and Lee {\em et al.} (2019) are not compatible with missing values in original data. BOISE, however, as a Bayes decision framework, can naturally handle with missing values in original data $x_0$. We will briefly discuss about missing data imputation in this section.

There are two procedures in BOISE that require missing data imputation: we need imputed data to update cluster labels in DPMM clustering, and we need a complete $x_0$ to determine $a_{kj}$ and $b_{kj}$ in informer selection. Therefore, we have two different strategies for missing value imputation. For DPMM clustering, we impute all missing entries after a thorough update of labels, that is, after we update all cluster labels $c_i$ for $i=1,\cdots,m$, we re-sample the value of missing entries based on posterior distribution $H_{-i,c}(\phi) = P\left(\phi\,|\,G_0,\{x_j:j\neq i, c_j = c\} \right)$. When all the clustering assignment samples are drawn, we summarize how many times each missing entry is imputed as $1$, say $N_{ij}$, and impute $x_{ij}$ as $1$ if $N_{ij}$ is greater than some threshold, like $50\%$ of clustering sample size. Then we can proceed with this final imputed data. The procedure of missing data imputation is in Algorithm $5$:

\begin{algorithm}
\caption{Missing data imputation}\label{Algo5}
\begin{algorithmic}[1]
\State Set the prior mass $m_0$ and other hyperparameters;
\State Set sample size of clustering assignments $M$, and gaps between successive draws $N$
\State \textbf{Initialize}: Let $N_{ij} = 0$ for each missing entry $x_{ij}$. Sample $c_1,\cdots,c_m$ with regard to Chinese restaurant process CR($m_0$);
\While{Sample size $< M$}
    \State For each missing entry $x_{ij}$ currently in cluster $c$, sample a success rate $\phi_{cj}$ from $P\left(\phi\,|\,G_0,\{x_j:j\neq i, c_j = c\} \right)$.
    \State Impute $x_{ij}$ with ${\rm Bernoulli}(\phi_{cj})$. Let $N_{ij} \gets N_{ij}+x_{ij}$
    \For{$i=1, \cdots, m$}
        \State Update $i$th cluster label from posterior distribution $P(c_i\,|\,c_{-i},y_i)$ in~(\ref{eq:Postc})
    \EndFor
\State Sample one clustering assignment after every $N$ label updates.
\EndWhile
\If {$N_{ij} > M/2$}
    \State Impute $x_{ij}$ as $1$.
\Else
    \State Impute $x_{ij}$ as $0$.
\EndIf
\end{algorithmic}
\end{algorithm}

@@

However, these experimentally determined thresholds contain information from non-informer compounds and should be kept away from all methods. In order to binarize the inhibition values to get the intermediate data $x_A$ for BOISE and Accelerated-BOISE, we follow the same procedure as Coding Selection (CS) and Adaptive Selection (AS) did in Zhang, Ericksen, and Lee {\em et al.} (2019). That is, we focus on the sub-matrix of inhibition data $Z$ on informer compounds, say $Z_A$, and keep track of all the possible binary outcomes $x_l$ of each row of $Z_A$. For rows with same binary outcome $x_l$, we compute a centroid corresponding to this outcome, say $c_l$, then the binary outcome $x_A$ of new target with inhibition values $z_A$ is defined as:
\begin{align}
    \label{eq:Code}
    \begin{split}
        x_A &= x_{l^*} \\
        s.t.\,~ ||z_A - c_{l^*}|| &= \argmin\limits_{l} ||z_A-c_l||
    \end{split}
\end{align}
Procedure~(\ref{eq:Code}) is exactly the same as CS and AS. There may be better ways to derive intermediate data, however, we follow the same procedure since we want to show our better performance comes from BOISE and Accelerated-BOISE themselves but not from a more accurate intermediate data. The result is summarized in following Table 1:
\bigskip
\begin{table}[ht!]
\centering
\caption{\label{tbl:t1}
{\bf Recovery counts by different informer based ranking methods on new kinase targets PknB, BGLF4 and ROP18 using PKIS1 data set.}}

@@

\subsubsection{Retrospective Results}
We conduct retrospective leave one out cross validation for each one of $224$ PKIS1 targets. For BOISE, we use the informer size $n_A = 8$ as it is quite computational intensive. For Accelerated-BOISE, both $n_A= 8$ and $n_A = 16$ are used to compare different methods. Results from $224$ separate experiments are summarized in both tables and graphs.

When informer set size $n_A = 8$, Figure 1 summarize cross validation results with respect to NEF10 metric, and Figure 2 summarize the results for ROCAUC metric. The summary statistics are in Table 2. From the graphs we can see that BOISE and Accelerated-BOISE are both significantly better than other baseline and non-baseline methods, and BOISE performs similarly to Accelerated-BOISE. Although BOISE is still slightly better than Accelerated-BOISE overall, as it is optimal theoretically, their difference is not statically significant.

>>>>

\usepackage{amssymb,amsmath,amsfonts,bbm,eurosym,geometry,ulem,graphicx,caption,color,setspace,sectsty,comment,footmisc,caption,natbib,pdflscape,subfigure,array}
\usepackage{algorithm, algorithmicx}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{appendix}
\usepackage{xcolor}
\usepackage{hyperref}

@@

\begin{titlepage}
\title{Bayes Optimal Informer Sets for Drug Discovery\thanks{abc}}
\author{Peng Yu, {\sc contributing members of UW drug discovery group}, \\  \& Michael A. Newton\thanks{abc}}
\date{\today}
\maketitle
\begin{abstract}
In early stage drug discovery, there has been significant recent effort in developing compound prioritization methods for a new target from minimal data. One kind of such methods is informer based ranking (IBR) methods, which select an initial subset of compounds, so-called informer set, to be screened based on original data, and rank additional compounds for the new target based on this additional screening data and original data. However, the selection of informer set in previous methods is often heuristic, based on chemical diversity or randomized algorithms with no statistical guarantees. In this article, we  formalize the informer set ranking procedure as a two-stage decision problem and introduce a new method named as Bayes Optimal Informer SEt (BOISE) to select informer set as well as ranking the compounds against the new target. BOISE is a fully Bayesian procedure that selects informer set based on posterior expected loss and ranks compounds based on posterior expectation. It is guaranteed to be Bayes rule under proposed model and loss if computed exactly, and we propose a computationally efficient version of BOISE that achieves a balance between ranking performance and computation complexity. Both proposed methods and previous methods are applied on human kinase chemogenomics data set and anti-cancer drug sensitivity data set and evaluated under various ranking criteria. We demonstrate that proposed methods are significantly better than previous methods under all metrics. 
\vspace{0in}\\
\noindent\textbf{Keywords:} key1, key2, key3\\
\vspace{0in}\\
\noindent\textbf{JEL Codes:} key1, key2, key3\\

@@

\section{Introduction} \label{sec:introduction}
Informer based ranking (IBR) strategies are a family of methods that arise from early-stage drug discovery (Zhang, Ericksen, and Lee {\em et al.} 2019). It is designed for computational guidance of drug-discovery experiments. The context is virtual screening (Sliwoski {\em et al.} 2014).  Briefly speaking, a chemical screening facility receives material representing a purified protein target, and is tasked with identifying drugs that may have some inhibitory effect on the bioactivity of the protein.  For example, Zhang, Ericksen, and Lee {\em et al.} (2019) report on drugs that affect certain protein kinases.
The facility has specific biochemical assays that test for inhibitory effects of a given drug, and it has access to a very large library of different drug compounds, some small fraction of which are expected to have bioactivity against the new target.   
Costs limit the number of compounds that may be screened experimentally.   The virtual screening task is to prioritize the drugs in advance of experimental screening by some computations so as to
improve the hit rate of the experiments performed on the new target.

Virtual screening may operate on different kinds of target and compound data, including possibly chemical structures or the results of experimental assays.   For example, when chemical structures are available it is possible to use sophisticated molecular docking programs that aim to find best fits between target and compound structures (e.g. Kitchen {\em et al.} 2004, Erickson {\em et al.} 2017)
We focus on informer based ranking strategies that need no information on chemical structure, only rely on bioactivity data of many compounds against many targets, and not any bioactivity data from the new target on test.  With these initial bioactivity data,   informer based ranking method
identifies a subset of compounds -- the {\em informer set} -- which
represent a small fraction of the available compounds and which we
will test experimentally for bioactivity against the new target.  Subsequently, we use the full set of data (initial data on other targets
for many compounds and intermediate data on informer compounds on 
the new target) in order to prioritize the remaining, non-informer
compounds for their likely bioactivity against the new target.  If
we have done a good job in building the informer set, then the 
top of the prioritized list of non-informers will be enriched
for truly active compounds.  We will have spent computational time
building the informer set from initial data, and experimental resources
obtaining intermediate data; the payoff is that further experiments
will return truly active drug compounds at  a higher rate than
if we had selected compounds to test by some other scheme.

One common misunderstanding in informer based ranking method is that someone will tend to select the most frequent hitters in bioactivity matrix as the informer set. We come up with a simple counterexample in Appendix to show that informer set ranking problem is a more complicated problem. In short, most informative compounds are not those that are most likely to be active. Zhang, Erichsen and Lee {\em et al.} (2019) introduced several informer based ranking strategies and compared
them empirically in data on protein kinases. Though differing in detail,
the proposed strategies all partition the initial
bioactivity data, producing  clusters of relatively similar targets; 
then informer compounds are  those that are predictive of the cluster label of a target.  Among kinase targets, the proposed bioactivity-based strategies had better predictive
performance than commonly used strategies based upon chemical
structure.   Since effective informer based ranking strategies leverage statistical
patterns in the initial bioactivity data, we reason that
statistical modeling may provide a useful approach to deriving
more effective IBR strategies than are currently available.

In this article, we formalize the informer set ranking procedure as a two-stage decision problem and introduce a new method named as Bayes Optimal Informer SEt (BOISE) to select informer set as well as ranking the compounds against the new target.  The proposed method is fully Bayesian and can be applied to other ranking problems where only limited information can be derived from an unknown new target. We show in our analysis that this method is guaranteed to be Bayes rule if computed exactly, and can be efficiently computed with an approximate version to achieves a balance between ranking performance and computation complexity.  The proposed methods are applied to human  kinase  chemogenomics  data  set PKIS1 and anti-cancer drug sensitivity data set GDSC1. We compare new methods with previous methods under different informer set sizes and different ranking metrics, such as NEF10, ROCAUC, MCC and F1. We demonstrate that new methods are significantly better than previous methods under these data sets and metrics.

\section{Methodology}

@@

We are given an $m \times n$ binary matrix $x_0$ containing bioactivity data measured on $m$ targets with index set $I = \{1,2,\cdots,m\}$, and $n$ compounds with index set $J=\{1,2,\cdots,n\}$.  Each entry $x_{ij}$ of $x_0$ is a summary of experiments, such as inhibition assays,
of compound $j$ on target $i$. To be more specific, $x_{ij} = 1$ means compound $j$ is active on target $i$ while $x_{ij} = 0$ means inactive. Binary outcome is of more interest to us because our task is to rank all active compounds on a given target prioritized above the inactives. Although our method is based on binary matrix, it can be applied to any continuous bioactivity matrices as we can always convert a continuous matrix into a binary matrix. For example, if $z$ is a continuous matrix, we can use a target-wise threshold criterion (\ref{eq:thresh}) to binarize $z$ based on sample mean and standard deviation of each row:

\begin{eqnarray}\label{eq:thresh}
x_{ij} = \begin{cases}
		1, &\text{ if }
	      z_{ij} \geq \mbox{mean}[z_{i\cdot}] + 2 \times \mbox{s.d.}[z_{i\cdot}],\\
		0 & \text{ if }
	      z_{ij} < \mbox{mean}[z_{i\cdot}] + 2 \times \mbox{s.d.}[z_{i\cdot}].
	\end{cases}
\end{eqnarray}
2SD threshold (\ref{eq:thresh}) is widely used to binarize outcomes in biomedical science.

@@

In Zhang, Ericksen, and Lee {\em et al.} (2019), there are $3$ novel informer-based-ranking (IBR) methods: Regression Selection (RS), Coding Selection (CS), Adaptive Selection (AS), as well as $2$ baseline methods: Baseline Chemometric (BC) and Baseline Frequent-hitters (BF). In this paper, we will introduce a new IBR method named as Bayes Optimal Informer SEt (BOISE) that is optimal in the context of Bayes decision theory.

@@

\textbf{Note:} The reason we use notation $\mathcal{C}$ is that we consider clustering structure as a random process on all possible partitions of countable targets space. Dirichlet process mixture model (DPMM) is a classical method to handle this kind of clustering structures, and we are using a special case of DPMM called Chinese restaurant (CR) process based on some prior mass
$m_0$ to make the posterior inference. There are two main advantages of using DPMM instead of traditional clustering methods like spectral clustering, hierarchical clustering and k-means: On the one hand, we don't need to specify the number of clusters $K$ beforehand, or manually cut the tree to get the clustering structure. All we need to set is a prior mass $m_0$ that describes our belief of divergence and all the rest is data-driven. Thus, DPMM is less subjective than other methods. On the other hand, DPMM is a non-parametric Bayesian clustering method and hence is compatible with our Bayesian framework. Also, since the sample size is not very large($\approx 200$ targets), number of targets falling in each cluster can be unbalanced, with some clusters containing more than $50$ targets while other clusters containing less than 5 targets. DPMM is well designed to address this unbalanced impact under Bayesian framework. We will come to details of DPMM later in computation section.

@@

We are faced with a two-stage decision problem: First, we need a rule to select an informer set $A$ of size $n_A$ from $x_0$. Then we need to select a top set $T$ of size $n_T$ based on initial data $x_0$, informer set $A$ and intermediate data $x_A$. Denote these decision rules as $A = A(x_0)$ and $T = T(A,x_A,x_0)$ to show their dependency.

@@

Consider when we have initial data $x_0$ as well as intermediate data $x_A$ on
some pre-specified informer set $A$, we can get the posterior distribution of $\theta$:
\begin{eqnarray*}
p(\theta|x_0,x_A) = \frac{ p(x_0|\theta) p(x_A|x_0, \theta) p(\theta) }{ p(x_0, x_A) }.
\end{eqnarray*}
For all compounds $j$, we then construct the posterior means under this distribution as:
\begin{eqnarray*}
\hat \theta_{i^*j} = \mathbb{E}\left( \theta_{i^*j} |x_A, x_0 \right)
\end{eqnarray*}
We define a particular top set rule $T^*$ to select the top $n_T$ compounds with largest posterior mean:
\begin{eqnarray}
\label{eq:topset}
T^*(A,x_A,x_0) = \left\{ j \in J: \; \sum_{k\in J} \mathbbm{1}(\hat \theta_{i^*k} > \hat \theta_{i^*j}) \leq n_T \right\}.
\end{eqnarray}
For the loss~(\ref{eq:loss}), it is known that $T^*$ minimizes the Bayes risk (Henderson 2016).
We utilize this finding to resolve one part of our compound-decision problem, obtaining:

@@

The first method is a straightforward sequential selection. At first we evaluate ${\rm PEL}_1(x_0,A)$ for all $A=1,\cdots,n$ to select the best $A$ with least ${\rm PEL}_1$ loss. Then at each iteration we memorize all previous decisions $A$ and select one best element $j\in J\setminus A$ that minimizes ${\rm PEL}_1(A\cup \{j\})$. The procedure can be summarized as following Algorithm \ref{Algo3}:

\begin{algorithm}
\caption{Exact Informer Selection}\label{Algo3}
\hspace*{\algorithmicindent} \textbf{Input:} Initial data $x_0$, size of informer set $n_A$, size of top set $n_T$.\\
\hspace*{\algorithmicindent} \textbf{Output:} Selected informer set of length $n_A$
\begin{algorithmic}[1]
\State \textbf{Initialization}: Evaluate ${\rm PEL}_1(x_0,A)$ for all $|A|=1$. Let $A^* = \argmin_{|A|=1}{\rm PEL}_1(x_0,A) $.
\While{$|A^*| < n_A$}
    \State Evaluate ${\rm PEL}_1(x_0,A^*\cup\{j\})$ for each $j\in J\setminus A^*$.
    \State Let $A^* \gets A^*\cup\{j^*\}$, where $j^* = \argmin_{j}{\rm PEL}_1(x_0,A^*\cup\{j\})$.
\EndWhile
\State Return $A^*$ as selected informer set.
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Accelerated Informer Selection}\label{Algo4}
\hspace*{\algorithmicindent} \textbf{Input:} Initial data $x_0$, size of informer set $n_A$, size of top set $n_T$.\\
\hspace*{\algorithmicindent} \textbf{Output:} Selected informer set of length $n_A$
\begin{algorithmic}[1]
\State \textbf{Initialization}: Let $A^* = \{\}$. Sample clustering assignments $\mathbb C$ from $p(\mathcal C\,|\, x_0)$ with Algorithm \ref{Algo1}.
\For{each $\mathcal C$ in $\mathbb C$}
    \State Sample $\mathbb{X}(\mathcal{C})$ from $p\left(x_{i^*}\,|\, \mathcal C,x_0\right)$ as in~(\ref{eq:Postphi}) and~(\ref{eq:Postc}).
    \Comment{\textbf{Auxiliary sample}}
\EndFor
\While{$|A^*| < n_A$}
    \For{each $j\in J\setminus A^*$}
        \For{each $\mathcal{C}$ in $\mathbb C$ and corresponding $\mathbb{X}(\mathcal{C})$}
            \State Evaluate $H(c(i^*)\,|\,\mathcal{C},x_0,x_{A^*\cup j})$ as in~(\ref{eq:Entropy});
        \EndFor
        \State Average over $x_{A^*\cup j}$ and $\mathcal{C}$ to get $\mathcal H(A^*\cup\{j\} )$ in~(\ref{eq:Score}). 
    \EndFor
    \State Let $A^*\gets A^*\cup \{j^*\}$, where $j^*$ is the minimizer of $\mathcal H(A^*\cup\{j\})$.
\EndWhile
\State Return $A^*$ as selected informer set.
\end{algorithmic}
\end{algorithm}
As mentioned in the beginning of this section, the major advantage of Algorithm \ref{Algo4} compared with Algorithm \ref{Algo3} is running time. In Algorithm \ref{Algo4} we bypass the computation of $\mathbb E\left(\theta_{ij}\,|\, x_0,x_A\right)$ and only use $p_k = P(i^*\in c_k\,|\, \mathcal C,x_0,x_A)$ to select informer set. This can improve our BOISE running time from days to hours while still maintaining similar performance. Therefore, we name it as "Accelerated-BOISE" in our empirical analysis. Also, although Accelerated-BOISE uses different informer selection method as BOISE does, they both share the same top set rule $T^*(A,x_A,x_0)$, which scores each $j\in J$ with their posterior expectation and choose the top $n_T$ as top set.

\subsection{Missing Data Imputation}
As a Bayesian computation procedure, BOISE is naturally compatible with missing data imputations while previous informer based ranking methods are only designed for complete data sets. We will briefly discuss about missing data imputation in this section. 
There are two procedures in BOISE that require missing data imputation: we need imputed data to update cluster labels in DPMM clustering, and we need a complete $x_0$ to determine $a_{kj}$ and $b_{kj}$ in informer selection. Therefore, we have two different strategies for missing value imputation summarized in Algorithm \ref{Algo5}:
\begin{algorithm}
\caption{Missing data imputation}\label{Algo5}
\begin{algorithmic}[1]
\State Set the prior mass $m_0$ and other hyperparameters;
\State Set sample size of clustering assignments $M$, and gaps between successive draws $N$
\State \textbf{Initialize}: Let $N_{ij} = 0$ for each missing entry $x_{ij}$. Sample $c_1,\cdots,c_m$ with regard to Chinese restaurant process CR($m_0$);
\While{Sample size $< M$}
    \State For each missing entry $x_{ij}$ currently in cluster $c$, sample a success rate $\phi_{cj}$ from $P\left(\phi\,|\,G_0,\{x_j:j\neq i, c_j = c\} \right)$.
    \State Impute $x_{ij}$ with ${\rm Bernoulli}(\phi_{cj})$. Let $N_{ij} \gets N_{ij}+x_{ij}$
    \For{$i=1, \cdots, m$}
        \State Update $i$th cluster label from posterior distribution $P(c_i\,|\,c_{-i},y_i)$ in~(\ref{eq:Postc})
    \EndFor
\State Sample one clustering assignment after every $N$ label updates.
\EndWhile
\If {$N_{ij} > M/2$}
    \State Impute $x_{ij}$ as $1$.
\Else
    \State Impute $x_{ij}$ as $0$.
\EndIf
\end{algorithmic}
\end{algorithm}

 For DPMM clustering, we impute all missing entries after a thorough update of labels, that is, after we update all cluster labels $c_i$ for $i=1,\cdots,m$, we re-sample the value of missing entries based on posterior distribution $H_{-i,c}(\phi) = P\left(\phi\,|\,G_0,\{x_j:j\neq i, c_j = c\} \right)$. When all the clustering assignment samples are drawn, we summarize how many times each missing entry is imputed as $1$, say $N_{ij}$, and impute $x_{ij}$ as $1$ if $N_{ij}$ is greater than some threshold, like $50\%$ of clustering sample size. Then we can proceed with this imputed complete data.

@@

However, these experimentally determined thresholds contain information from non-informer compounds and should be kept away from all methods. In order to binarize the inhibition values to get the intermediate data $x_A$ for BOISE and Accelerated-BOISE, we focus on the sub-matrix of inhibition data $Z$ on informer compounds, say $Z_A$, and keep track of all the possible binary outcomes $x_l$ of each row of $Z_A$. For rows with same binary outcome $x_l$, we compute a centroid corresponding to this outcome, say $c_l$, then the binary outcome $x_A$ of new target with inhibition values $z_A$ is defined as:
\begin{align}
    \label{eq:Code}
    \begin{split}
        x_A &= x_{l^*} \\
        s.t.\,~ ||z_A - c_{l^*}|| &= \argmin\limits_{l} ||z_A-c_l||
    \end{split}
\end{align}
Procedure~(\ref{eq:Code}) is exactly the same as CS and AS. There may be better ways to derive intermediate data, however, we follow the same procedure since we want to show our better performance comes from BOISE and Accelerated-BOISE themselves but not from a more accurate intermediate data. The result is summarized in following Table 1:
\bigskip
\begin{table}[ht!]
\centering
\caption{\label{tbl:t1}
{\bf Recovery counts by different informer based ranking methods on new kinase targets PknB, BGLF4 and ROP18 using PKIS1 data set.}}

@@

\subsubsection{Retrospective Results}
We conduct retrospective leave one out cross validation for each one of $224$ PKIS1 targets. For BOISE, we use the informer size $n_A = 8$ as it is quite computational intensive. For Accelerated-BOISE, both $n_A= 8$ and $n_A = 16$ are used to compare different methods. Results from $224$ separate experiments are summarized in violin plots.

When informer set size $n_A = 8$, Figure \ref{fig:pkis8} summarize cross validation results with respect to NEF10 and ROCAUC metrics. From pairwise tests, we can see that BOISE and Accelerated-BOISE are both significantly better than other baseline and non-baseline methods, and BOISE performs similarly to Accelerated-BOISE. Although BOISE is still slightly better than Accelerated-BOISE overall, as it is optimal theoretically, their difference is not statically significant. 
To be more specific, it can be seen from Figure \ref{fig:pkis8} that Adaptive Selection (AS8) is the best method other than (AC-)BOISE8 under NEF10 metric. From paired t-test between AS8 and (AC-)BOISE8, the mean difference between BOISE8 and AS8 is $0.0608$, with p-value $<2.2\times 10^{-16}$; the mean difference between AC-BOISE8 and AS8 is $0.0520$, with p-value $=6.466\times 10^{-14}$. Similarly, Coding Selection (CS8) is the best method other than (AC-)BOISE8 under ROCAUC. The mean difference between BOISE8 and CS8 is $0.0283$, with p-value $=3.73\times 10^{-6}$ and the mean difference between AC-BOISE8 and CS8 is $0.0258$, with p-value $=3.035\times 10^{-5}$.
